### 2024-01-20 outputs/paper/paper.qmd
Graph Update
- Graph Updated
- Added Reference

---
title: "Police Arrests and Ethnic Background"
author: "Aviral Bhardwaj"
date: today
date-format: long
number-sections: true
format: pdf
bibliography: references.bib
---

```{r}
#| echo: false
# All the required packages
## Installing Packages only needs to be done once

#install.packages("tidyverse")
#install.packages("tinytex")
#install.packages("opendatatoronto")
# install.packages("kableExtra")
```

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(opendatatoronto)
library(tinytex)
library(knitr)
library(dplyr)
library(ggplot2)
library(janitor)
```

```{r}
#| warning: false
#| message: false
#| include: false

# Get package data

# get package
package <- show_package("police-race-and-identity-based-data-collection-arrests-strip-searches")
package

# get all resources for this package
resources <- list_package_resources("police-race-and-identity-based-data-collection-arrests-strip-searches")

# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))

# load the first datastore resource as a sample
data <- filter(datastore_resources, row_number()==1) %>% get_resource()

# Save the data in CSV
write.csv(data, "../data/raw_data.csv")
```

## Introduction

## Data 
This data is sourced from the Toronto Open Data Portal `opendatatoronto` [@rOpenDataToronto]. The following data is utilized for a more in-depth analysis of arrests and strip searches involving various ethnic and identity-based groups in Toronto. The information is gathered by the Toronto Police Service (TPS) and is collected under the authority of the Police Services Act. Its purpose is to enhance comprehension of the relationship between the police and the community. The data covers the period from 2020 to 2021.

The data underwent cleaning and analysis using the R programming language [@r]. Cleaning was performed with the `tidyverse` package [@rTidyverse], involving the removal of unnecessary columns for the analysis. Subsequently, analysis was conducted utilizing the `dplyr` package [@rDplyr], with the data grouped by ethnic and identity-based categories. Visualization was then executed using the `ggplot2` package [@rGgplot2], depicting the number of arrests and strip searches across various ethnic and identity-based groups.

The raw data included 32,000 arrests data with 26 variables of analyzable data, that included variables such as Arrest year, Arrest_Month, Perceived Race, Age Group, Youth at Arrest and Booked which we will be using for the paper.

```{r}
#| warning: false
#| message: false
#| include: false
# Adjusting and cleaning the name of the variable
cleaned_data <- data %>% 
  clean_names() %>% 
  clean_names(case = "snake")

# Isolating the columns that are needed for the analysis
cleaned_data <- cleaned_data |>
  select(arrest_year, arrest_month, perceived_race, age_group_at_arrest, youth_at_arrest_under_18_years, booked)

# Renaming the columns
cleaned_data <- cleaned_data |>
  rename(youth = youth_at_arrest_under_18_years, age_group = age_group_at_arrest)

# Saving the cleaned data
write.csv(cleaned_data, "../data/cleaned_data.csv")
```

```{r}
#| warning: false
#| message: false
#| include: false
# Make a table of the data with the number of arrests and strip searches by ethinic and identity based groups for the year 2020, and 2021
cleaned_data <- read.csv("../data/cleaned_data.csv")

# Mutating the date for age groups involving Aged 17 years and younger and Aged 17 years and under essentially they mean the same thing
cleaned_data <- cleaned_data %>%
  mutate(age_group = case_when(
    age_group == "Aged 17 years and younger" ~ "Under 18 years",
    age_group == "Aged 17 years and under" ~ "Under 18 years",
    TRUE ~ age_group
  ))

# List of the ethnic and identity based groups that are in the data
# ethnic_backgrounds <- cleaned_data |> select(perceived_race) |> unique()

# List of the years that are in the data
# years <- cleaned_data |> select(arrest_year) |> unique()

# List of the age groups that are in the data
# age_groups <- cleaned_data |> select(age_group) |> unique()

write.csv(cleaned_data, "../data/cleaned_data.csv")
```

```{r}
#| warning: false
#| message: false
#| echo: false
# Generate a table of the data with the number of arrests and strip searches by ethinic and identity based groups for the year 2020, and 2021
cleaned_data <- read.csv("../data/cleaned_data.csv")

table1 <- cleaned_data |> 
  group_by(arrest_year, perceived_race) |> 
  summarise(n = n(), .groups = "drop") |> 
  arrange(desc(n)) |> 
  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)

kable(table1,
      caption = "Number of arrests by ethinic background for the year 2020, and 2021",
      align = "c",
      col.names = c("Ethinic and Identity Based Groups", "2020", "2021"),
      booktabs = TRUE)

ggplot(
  table1, aes(x = `perceived_race`, y = `2020`, fill = `2021`)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Number of arrests by ethinic background for the year 2020, and 2021",
       x = "Ethinic and Identity Based Groups",
       y = "Number of Arrests") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_fill_gradient(low = "blue", high = "red") +
  theme(plot.title = element_text(hjust = 0.5))

```

```{r}

```

## References

### 2024-01-22 outputs/paper/paper.qmd
Additional Graphs
- Added Additional Graphs
- Fixed the styling
- Renamed the figures

@@
 # install.packages("kableExtra")
+# install.packages("patchwork")
 ```
@@
 library(janitor)
+library(patchwork)
 ```
@@
 #| include: false
+
 # Adjusting and cleaning the name of the variable
@@
 
-# Isolating the columns that are needed for the analysis
 cleaned_data <- cleaned_data |>
-  select(arrest_year, arrest_month, perceived_race, age_group_at_arrest, youth_at_arrest_under_18_years, booked)
+  select(
+    arrest_year, perceived_race, age_group_at_arrest, youth_at_arrest_under_18_years, booked, search_reason_cause_injury, search_reason_assist_escape, search_reason_possess_weapons, search_reason_possess_evidence)
 
-# Renaming the columns
-cleaned_data <- cleaned_data |>
-  rename(youth = youth_at_arrest_under_18_years, age_group = age_group_at_arrest)
+# Combining the columns that are similar
+cleaned_data <- cleaned_data %>%
+  mutate(search_reason = case_when(
+    search_reason_cause_injury == 1 ~ "1",
+    search_reason_assist_escape == 1 ~ "1",
+    search_reason_possess_weapons == 1 ~ "1",
+    search_reason_possess_evidence == 1 ~ "1",
+    TRUE ~ "0"
+  ))
 
-# Saving the cleaned data
-write.csv(cleaned_data, "../data/cleaned_data.csv")
-```
-
-```{r}
-#| warning: false
-#| message: false
-#| include: false
-# Make a table of the data with the number of arrests and strip searches by ethinic and identity based groups for the year 2020, and 2021
-cleaned_data <- read.csv("../data/cleaned_data.csv")
+# Rename the columns
+cleaned_data <- cleaned_data %>%
+  rename(
+    age_group = age_group_at_arrest,
+    youth_at_arrest_under_18_years = youth_at_arrest_under_18_years
+  )
 
@@
   ))
+
+# Mutating for booked with NA values
+cleaned_data <- cleaned_data %>%
+  mutate(booked = case_when(
+    is.na(booked) ~ "0",
+    booked == 1 ~ "1",
+    TRUE ~ "Not Booked"
+  ))
+
+cleaned_data <- cleaned_data |>
+  select(
+    arrest_year, perceived_race, age_group, youth_at_arrest_under_18_years, booked, search_reason)
+
+write.csv(cleaned_data, "../data/cleaned_data.csv")
+
+```
+
+```{r}
+#| warning: false
+#| message: false
+#| include: false
+# Make a table of the data with the number of arrests and strip searches by ethinic and identity based groups for the year 2020, and 2021
+# cleaned_data <- read.csv("../data/cleaned_data.csv")
 
@@
 write.csv(cleaned_data, "../data/cleaned_data.csv")
+```
+
+```{r}
+cleaned_data <- read.csv("../data/cleaned_data.csv")
+
+table <- cleaned_data |>
+  group_by(arrest_year, perceived_race, search_reason) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  arrange(desc(perceived_race)) %>%
+  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)
+
+table <- pivot_longer(table, cols = c("2020", "2021"), names_to = "year", values_to = "value")
+
+ggplot(table, aes(x = perceived_race, y = value, fill = factor(search_reason))) +
+  geom_bar(stat = "identity", position = "fill") +
+  facet_wrap(~ year, scales = "free_y") +
+  labs(title = "Percent Stacked Bar Chart Based on 'booked'",
+       x = "Perceived Race",
+       y = "Percentage",
+       fill = "Search Reason") +
+  theme_minimal() +
+  scale_fill_discrete(labels = c("Not Searched", "Searched")) +
+  theme(legend.position = "right") + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
 ```
@@
 
-ggplot(
-  table1, aes(x = `perceived_race`, y = `2020`, fill = `2021`)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  labs(title = "Number of arrests by ethinic background for the year 2020, and 2021",
-       x = "Ethinic and Identity Based Groups",
-       y = "Number of Arrests") +
+table1_long <- pivot_longer(table1, cols = c("2020", "2021"), names_to = "year", values_to = "value")
+```
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+table2 <- cleaned_data |>
+  filter(booked == 1) |>
+  group_by(arrest_year, perceived_race) |>
+  summarise(n = n(), .groups = "drop") |>
+  arrange(desc(n)) |>
+  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)
+
+kable(table2,
+      caption = "Arrests Booked by ethnic background for the year 2020, and 2021",
+      align = "c",
+      col.names = c("Ethinic and Identity Based Groups", "2020", "2021"),
+      booktabs = TRUE)
+
+table2_long <- pivot_longer(table2, cols = c("2020", "2021"), names_to = "year", values_to = "value")
+```
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+cleaned_data <- read.csv("../data/cleaned_data.csv")
+
+cleaned_data <- cleaned_data |>
+  select(
+    arrest_year,
+    perceived_race,
+    booked
+  )
+
+# Group 
+cleaned_data <- cleaned_data %>%
+  group_by(arrest_year, perceived_race, booked) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  arrange(desc(perceived_race)) %>%
+  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)
+
+cleaned_data <- pivot_longer(cleaned_data, cols = c("2020", "2021"), names_to = "year", values_to = "value")
+```
+
+```{r police_arrests, fig.cap = "Number of arrests by ethinic background for the year 2020, and 2021", fig.width=10, fig.height=6, fig.align='center'}
+ggplot(cleaned_data, aes(x = perceived_race, y = value, fill = factor(booked))) +
+  geom_bar(stat = "identity") +
+  facet_wrap(~ year, scales = "free_y") +
+  labs(title = "Stacked Bar Chart Based on 'booked'",
+       x = "Perceived Race",
+       y = "Value",
+       fill = "Booked") +
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  scale_fill_gradient(low = "blue", high = "red") +
-  theme(plot.title = element_text(hjust = 0.5))
+  scale_fill_discrete(labels = c("Not Booked", "Booked")) +
+  theme(legend.position = "right") + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
@@
 ```{r}
+#| warning: false
+#| message: false
+#| echo: false
+cleaned_data <- read.csv("../data/cleaned_data.csv")
+cleaned_data <- cleaned_data |>
+  select(
+    arrest_year,
+    perceived_race,
+    age_group
+  )
+
+# Group 
+cleaned_data <- cleaned_data %>%
+  group_by(arrest_year, perceived_race, age_group) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  arrange(desc(perceived_race)) %>%
+  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)
+
+cleaned_data <- pivot_longer(cleaned_data, cols = c("2020", "2021"), names_to = "year"  ,values_to = "value")
+```
+
+```{r fig2, fig.cap = "Number of arrests by ethinic background for the year 2020, and 2021", fig.width=10, fig.height=6, fig.align='center'}
+#| warning: false
+#| message: false
+#| echo: false
+#| eval: true
+# Dot Plot
+ggplot(cleaned_data, aes(x = perceived_race, y = value, fill = factor(age_group))) +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 0.7, binwidth = 80) +
+  facet_wrap(~ year, scales = "free_y") +
+  labs(title = "Stacked Bar Chart Based on 'booked'",
+       x = "Perceived Race",
+       y = "Value",
+       fill = "Booked") +
+  theme_minimal() +
+  theme(legend.position = "bottom") + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
+
+```{r fig3, fig.cap = "Number of arrests by ethinic background for the year 2020, and 2021", fig.width=10, fig.height=6, fig.align='center'}
+#| warning: false
+#| message: false
+#| echo: false
+#| eval: true
+# Bar plot
+ggplot(cleaned_data, aes(x = perceived_race, y = value, fill = factor(age_group))) +
+  geom_bar(stat = "identity") +
+  facet_wrap(~ year, scales = "free_y") +
+  labs(title = "Stacked Bar Chart Based on 'booked'",
+       x = "Perceived Race",
+       y = "Value",
+       fill = "Booked") +
+  theme_minimal() +
+  theme(legend.position = "right") + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 

### 2024-01-23 outputs/paper/paper.qmd
Commit Backup
Commit Backup

--- 2024-01-20 outputs/paper/paper.qmd
@@
 ---
-title: "Police Arrests and Ethnic Background"
+title: "No Evidence of Bias in Arrests Made by the Toronto Police Service"
 author: "Aviral Bhardwaj"
+abstract: The ethnic identity of citizens arrested by the police has consistently been a topic of discussion, often highlighting the pronounced skewness of certain ethnic identities.
+This paper seeks to analyze data from the Toronto Open Data Portal to investigate whether there is any bias in arrests made by the Toronto Police Service between 2020 and 2021, while accounting for factors such as age, booked after arrest, and search on arrest.
+While the population distribution of various ethnic identities varies significantly, it was found that .....
+Further analysis can provide valuable insights for policymakers, law enforcement agencies, and community advocates in fostering a more equitable society while emphasizing the importance of promoting transparency and accountability in policing.
 date: today
@@
 library(janitor)
+library(patchwork)
+library(kableExtra)
 ```
@@
 
-The raw data included 32,000 arrests data with 26 variables of analyzable data, that included variables such as Arrest year, Arrest_Month, Perceived Race, Age Group, Youth at Arrest and Booked which we will be using for the paper.
+The raw dataset comprised 32,000 arrest records spanning both 2020 and 2021, encompassing 26 variables of analyzable data.
+However, for the analysis in this paper, we narrowed our focus to specific variables, including Arrest Year, Perceived Race, Age Group, Youth at Arrest, Booked, Search Reason Cause Injury, Search Reason Assist Escape, Search Reason Possess Weapons, and Search Reason Possess Evidence.
+Upon further examination the raw data contained values such as 'NA' under the Booked column and 'None' under the Search column.
+Using the `mutate` function [@rDplyr], we replaced the 'NA' and `None` values with 0 and 1 to summarize the data.
+Additionally, we consolidated similar columns, namely Search Reason Cause Injury, Search Reason Assist Escape, Search Reason Possess Weapons, and Search Reason Possess Evidence, into a unified column labeled "Search Reason." The process involved utilizing the `any()` function [@rDplyr], assigning a value of 1 to indicate the presence of any of these search motives.
 
@@
 #| include: false
+
+# Reading the data
+data <- read.csv("../data/raw_data.csv")
+
 # Adjusting and cleaning the name of the variable
@@
 
-write.csv(cleaned_data, "../data/cleaned_data.csv")
+# write.csv(cleaned_data, "../data/cleaned_data.csv")
 ```
+
+### Arrests by Ethnic and Identity-Based Groups for Year 2020 and 2021
+
+Table 1 (@tbl-table1) provides a summary of our data, revealing the number of individuals arrested from each prominent ethnic group in the years 2020 and 2021.
 
@@
 #| echo: false
-# Generate a table of the data with the number of arrests and strip searches by ethinic and identity based groups for the year 2020, and 2021
+#| label: tbl-table1
+#| tbl-cap: "Arrests by ethnic background for the year 2020, and 2021"
+## Table 1: Total Number of Arrests
 cleaned_data <- read.csv("../data/cleaned_data.csv")
@@
 kable(table1,
-      caption = "Number of arrests by ethinic background for the year 2020, and 2021",
       align = "c",
+      linesep = "",
       col.names = c("Ethinic and Identity Based Groups", "2020", "2021"),
       booktabs = TRUE)
+```
 
-ggplot(
-  table1, aes(x = `perceived_race`, y = `2020`, fill = `2021`)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  labs(title = "Number of arrests by ethinic background for the year 2020, and 2021",
-       x = "Ethinic and Identity Based Groups",
-       y = "Number of Arrests") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  scale_fill_gradient(low = "blue", high = "red") +
-  theme(plot.title = element_text(hjust = 0.5))
+### Booked after Arrest (Ethnic and Identity-Based Groups for Year 2020 and 2021)
+
+Table 2 presents a data summary indicating the count of individuals from prominent ethnic groups who were "booked" in criminal cases post-arrest in the years 2020 and 2021.
+Due to the extensive nature of the dataset, encompassing nine different age groups across eight ethnicities, summarizing the information for clarity posed a challenge.
+Consequently, we have opted to showcase a subset of the data specifically focusing on the age group of 25 to 34 years for improved readability.
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| tbl-cap: "Arrests Booked by ethnic background for the year 2020, and 2021"
+
+## Table 2: Booked After Arrest
+cleaned_data <- read.csv("../data/cleaned_data.csv")
+
+table2 <- cleaned_data |>
+  filter(booked == 1) |>
+  group_by(arrest_year, perceived_race) |>
+  summarise(n = n(), .groups = "drop") |>
+  arrange(desc(n)) |>
+  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)
+
+kable(table2,
+      align = "c",
+      linesep = "",
+      col.names = c("Ethinic and Identity Based Groups", "2020", "2021"),
+      booktabs = TRUE)
+```
+
+### Total number of arrests by age group for the year 2020 and 2021
+
+Table 3 presents a data summary indicating the count of different age groups of individuals from prominent ethnic groups in the years 2020 and 2021.
+Due to the extensive nature of the dataset, encompassing nine different age groups across eight ethnicities, summarizing the information for clarity posed a challenge.
+Consequently, we have opted to showcase a subset of the data specifically focusing on the age group of 25 to 34 years for improved readability.
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+
+## Table 3: Arrests by Age Group
+
+cleaned_data <- read.csv("../data/cleaned_data.csv")
+# 
+# # Group 
+table3 <- cleaned_data %>%
+  group_by(arrest_year, perceived_race, age_group) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  arrange(desc(perceived_race)) %>%
+  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)
+
+table3_small <- cleaned_data %>%
+  filter(age_group == "Aged 25 to 34 years") |>
+  group_by(arrest_year, perceived_race, age_group) %>%
+  summarise(n = n(), .groups = "drop") |>
+  arrange(desc(n)) |>
+  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)
+
+kable(table3_small,
+      caption = "Arrests Booked by ethnic background for the year 2020, and 2021",
+      align = "c",
+      linesep = "",
+      col.names = c("Ethinic and Identity Based Groups", "Age Group", "2020", "2021"),
+      booktabs = TRUE)
 
@@
 
+We would like to acknowledge that opendatatoronto [@rOpenDataToronto] provides the distribution of the population by ethnicity.
+However, due to the vast size of Toronto and its numerous wards, it has been challenging to compare all the data and cross-verify it with our crime rate data comprehensively.
+Nevertheless, a more detailed analysis can be conducted by comparing specific ethnic groups with each other.
+
+## Results
+Now that we have cleaned the data, we can begin to analyze the data.
+The following section will provide a summary of the data, and the trends that emerge from the data.
+Figure 1, Figure 2, Figure 3, Figure 4 was created using `ggplot` [@rGgplot2], displays the information that can be used as compare the trends as described above.
+
+### Number of Arrests in year 2020-2021
+With reference to Table 1 and the graph plotted the data shows a non-uniform distribution of arrests across ethnic groups, with the Black ethnic group and White group exhibiting the highest numbers of arrests in both 2020 and 2021.
+Furthermore, the data indicates a decrease in the number of arrests for the top two ethnic groups from 2020 to 2021.
+However, a contrasting trend emerges for other ethnic groups, where the number of arrests has increased over the same time period.
+
 ```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| eval: true
+#| fig.cap: "Number of arrests by ethinic background for the year 2020, and 2021"
+#| fig.width: 10
+#| fig.height: 6
+#| fig.align: 'center'
+#| fig.pos: 'htb'
+
+table1_long <- pivot_longer(table1, cols = c("2020", "2021"), names_to = "year", values_to = "value")
+
+# Create a plot
+ggplot(table1_long, aes(x = perceived_race, y = value, fill = year)) +
+  geom_bar(stat = "identity", position = 'dodge') +
+  labs(title = "Bar Chart for the number of arrests by ethinic background for the year 2020, and 2021",
+       x = "Ethnic and Identity Based Groups",
+       y = "Number of Arrests",
+       fill = "Year") +
+  theme_minimal() +
+  scale_fill_discrete(labels = c("2020", "2021")) +
+  theme(legend.position = "right") + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
+  scale_fill_brewer(palette = "Set1")
+```
+
+### Number of Arrests Booked in year 2020-2021
+Referring to Table 2 and the plotted graph, the data illustrates a consistent distribution of the number of individuals booked after arrests across various ethnic groups.
+Notably, there is a higher count of individuals booked after arrest compared to those who were arrested but not booked for a criminal case.
+Additionally, it is observed that, across all ethnic groups, the percentage of individuals booked falls within the range of 50% to 55%, and this pattern remains consistent in the year 2021.
+
+However, it is noteworthy that the Black ethnicity exhibits the highest number of individuals booked after arrest in both the years 2020 and 2021.
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| eval: true
+#| fig.cap: "Number of arrests booked by ethinic background for the year 2020, and 2021"
+#| fig.width: 10
+#| fig.height: 6
+#| fig.align: 'center'
+#| fig.pos: 'htb'
+
+table2_long <-
+  cleaned_data |>
+  select(
+    arrest_year,
+    perceived_race,
+    booked
+  ) |>
+  group_by(arrest_year, perceived_race, booked) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  arrange(desc(perceived_race)) %>%
+  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)
+
+table2_long <- pivot_longer(table2_long, cols = c("2020", "2021"), names_to = "year", values_to = "value")
+
+ggplot(table2_long, aes(x = perceived_race, y = value, fill = factor(booked))) +
+  geom_bar(stat = "identity", position = "fill") +
+  facet_wrap(~ year, scales = "free_y") +
+  labs(title = "Booked after Arrest in the year 2020 and 2021",
+       x = "Perceived Race",
+       y = "Percentage of Arrests Booked",
+       fill = "Status") +
+  theme_minimal() +
+  theme(legend.position = "right") + 
+  scale_fill_discrete(labels = c("Not Booked", "Booked")) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+### Number of Arrests by Age Group for the year 2020-2021
+As seen in Figure 3, the data indicates a non-uniform distribution of arrests across age groups, with the age group of 25 to 34 years and age group 35 to 44 exhibiting the highest number of arrests in both 2020 and 2021.
+Furthermore, the data indicates a decrease in the number of arrests for the top two age groups from 2020 to 2021.
+However, a contrasting trend emerges for other age groups, where the number of arrests has increased over the same time period.
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| eval: true
+#| fig.cap: "Number of arrests by Age-Group for the year 2020, and 2021"
+#| fig.width: 10
+#| fig.height: 6
+#| fig.align: 'center'
+#| fig.pos: 'htb'
+# Bar plot
+
+table3_long <- pivot_longer(table3, cols = c("2020", "2021"), names_to = "year", values_to = "value")
+
+ggplot(table3_long, aes(x = perceived_race, y = value, fill = factor(age_group))) +
+  geom_bar(stat = "identity") +
+  facet_wrap(~ year, scales = "free_y") +
+  labs(title = "Bar Chart for the number of arrests by Age-Group for the year 2020, and 2021",
+       x = "Perceived Race",
+       y = "Value",
+       fill = "Year") +
+  theme_minimal() +
+  theme(legend.position = "right") + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+table3_age_24_long <- pivot_longer(table3_small, cols = c("2020", "2021"), names_to = "year", values_to = "value")
+
+age_35_44 <- cleaned_data %>%
+  filter(age_group == "Aged 35 to 44 years") |>
+  group_by(arrest_year, perceived_race, age_group) %>%
+  summarise(n = n(), .groups = "drop") |>
+  arrange(desc(n)) |>
+  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0) |>
+  pivot_longer(cols = c("2020", "2021"), names_to = "year", values_to = "value")
+
+# Combining the data
+table3_age_24_long <- rbind(table3_age_24_long, age_35_44)
+
+ggplot(table3_age_24_long, aes(x = perceived_race, y = value, fill = factor(year))) +
+  geom_bar(stat = "identity", position = 'dodge') +
+  facet_wrap(~ age_group, scales = "free_y") +
+  labs(title = "Bar Chart for the number of arrests by Age-Group for the year 2020, and 2021",
+       x = "Perceived Race",
+       y = "Value",
+       fill = "Age Group") +
+  theme_minimal() +
+  theme(legend.position = "right") + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
@@
 
+## Discussion
 ## References

### 2024-01-23 outputs/paper/paper.qmd
Filter Updates
- Updated Cleaned Data
- Added Scatter Plot for Results

@@
 format: pdf
+toc: true
 bibliography: references.bib
@@
   mutate(age_group = case_when(
-    age_group == "Aged 17 years and younger" ~ "Under 18 years",
-    age_group == "Aged 17 years and under" ~ "Under 18 years",
+    age_group == "Aged 17 years and younger" ~ "Aged 17 years and under",
+    age_group == "Aged 17 years and under" ~ "Aged 17 years and under",
     TRUE ~ age_group
@@
   arrange(desc(perceived_race)) %>%
+  filter(!(perceived_race == "None")) %>%
   pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)
@@
 ### Number of Arrests by Age Group for the year 2020-2021
-As seen in Figure 3, the data indicates a non-uniform distribution of arrests across age groups, with the age group of 25 to 34 years and age group 35 to 44 exhibiting the highest number of arrests in both 2020 and 2021.
-Furthermore, the data indicates a decrease in the number of arrests for the top two age groups from 2020 to 2021.
-However, a contrasting trend emerges for other age groups, where the number of arrests has increased over the same time period.
+As depicted in Figure 3, the data reveals a non-uniform distribution of arrests across age groups, with the age groups of 25 to 34 years and 35 to 44 years exhibiting the highest number of arrests in both 2020 and 2021.
+Furthermore, there is a notable decrease in the number of arrests for these top two age groups from 2020 to 2021.
+In contrast, a different trend emerges for other age groups, where the number of arrests has increased over the same time period.
+
+Examining Figure 4, which illustrates the dot plot for the number of arrests by age group in the year 2020 and 2021 for the top two ethnic groups, "Black" and "White," it becomes evident that the age group 25 to 34 years represents the highest number of arrests for individuals of Black ethnicity.
+The trend for Black ethnicity indicates a decreasing curve from this age group onward.
+In contrast, for individuals of White ethnicity, the peak in arrests is observed in the age group 35 to 44 years.
+It is noteworthy that from age 45 to 65, the difference in the number of arrests between both ethnicities starts to converge.
 
@@
 
-table3_age_24_long <- pivot_longer(table3_small, cols = c("2020", "2021"), names_to = "year", values_to = "value")
-
-age_35_44 <- cleaned_data %>%
-  filter(age_group == "Aged 35 to 44 years") |>
-  group_by(arrest_year, perceived_race, age_group) %>%
-  summarise(n = n(), .groups = "drop") |>
-  arrange(desc(n)) |>
-  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0) |>
+table3_ii <- table3 %>%
+  filter(perceived_race %in% c("White", "Black")) %>%
+  filter(!(age_group == "None")) %>%
   pivot_longer(cols = c("2020", "2021"), names_to = "year", values_to = "value")
 
-# Combining the data
-table3_age_24_long <- rbind(table3_age_24_long, age_35_44)
-
-ggplot(table3_age_24_long, aes(x = perceived_race, y = value, fill = factor(year))) +
-  geom_bar(stat = "identity", position = 'dodge') +
-  facet_wrap(~ age_group, scales = "free_y") +
-  labs(title = "Bar Chart for the number of arrests by Age-Group for the year 2020, and 2021",
-       x = "Perceived Race",
+# Create a line Graph
+ggplot(table3_ii, aes(x = age_group, y = value, color = perceived_race)) +
+  geom_point() + 
+  geom_smooth(method = "loess", se = FALSE) +
+  facet_wrap(~ year, scales = "free_y") +
+  labs(title = "Scatter Plot for the number of arrests by Age-Group for the year 2021",
+       x = "Age Group",
        y = "Value",
-       fill = "Age Group") +
+       color = "Perceived Race") +  # Change fill to color
   theme_minimal() +
@@
 ```
+## Discussion
 
-## Discussion
 ## References

### 2024-01-23 outputs/paper/paper.qmd
Added Introduction
- Added Introduction to the Paper

@@
 ## Introduction
+In the vibrant metropolis of Toronto, the intersection between law enforcement practices and demographic characteristics holds profound implications for community dynamics and public policy.
+This paper undertakes a comprehensive analysis of arrest data spanning the years 2020 and 2021, seeking to unravel the intricate patterns that shape the city's legal landscape.
+
+Arrest data serves as a vital lens through which we can gain insights into the dynamics of law enforcement activities and their correlation with ethnicity and age.
+Our inquiry begins with a broad exploration of the distribution of arrests across various age groups, aiming to discern overarching trends and fluctuations.
+The age spectrum, a fundamental dimension of this analysis, reveals intriguing patterns that underscore the nuanced nature of arrests in Toronto.
+
+As we embark on this exploration, the data prompts critical questions about the socio-cultural, economic, and systemic factors that contribute to these observed patterns.
+This paper aims to contribute to a nuanced understanding of the dynamics of arrests in Toronto, providing insights that can inform targeted interventions, policy considerations, and community engagement efforts.
 

### 2024-01-23 outputs/paper/paper.qmd
Added Thanks
- Added thanks to github link

@@
 toc: true
+thanks: "Code and data supporting this analysis is available at: https://github.com/Aviral-03/Ethnic-Identity-Arrests-Toronto"
 bibliography: references.bib

### 2024-01-25 outputs/paper/paper.qmd
Scripts Update
- Added Scripts
- Added Simulation Data
- Added Tests
- Improved the introduction, data cleaning and results
- Removed extra sections from Data section

@@
 # install.packages("patchwork")
+# install.packages("janitor")
+# install.packages("gridExtra")
 ```
@@
 library(kableExtra)
+library(gridExtra)
 ```
@@
 In the vibrant metropolis of Toronto, the intersection between law enforcement practices and demographic characteristics holds profound implications for community dynamics and public policy.
-This paper undertakes a comprehensive analysis of arrest data spanning the years 2020 and 2021, seeking to unravel the intricate patterns that shape the city's legal landscape.
+Toronto, named Canada's most diverse city, encounters numerous challenges arising from the coexistence of ethnically diverse populations within the same neighborhoods.
+According to an article published by Georgetown University, the "ongoing debate surrounding police reforms is heavily influenced by historical connections with slavery and contemporary racial disparities" (@rJackson2022).
+This statement underscores the contentious nature of whether the race identity affiliation of police officers affects their policing practices, despite the expectation for fair and unbiased treatment.
+
+In a recent article published by the Toronto Police Services, it is highlighted that "Black, East/Southeast Asian, Middle Eastern, and Latino people were over-represented in reported use of force incidents compared to their presence in the enforcement action population" (@Fanfair2022).
+This indicates that, despite Toronto Police Services working towards a more equitable society, there are still hints of racial bias in the arrests made by the Toronto Police Services.Conversely this paper undertakes a comprehensive analysis of arrest data of various race and ethinic iden spanning the years 2020 and 2021 in the city of Toronto, seeking to unravel the intricate patterns that shape the city's legal landscape.
 
 Arrest data serves as a vital lens through which we can gain insights into the dynamics of law enforcement activities and their correlation with ethnicity and age.
-Our inquiry begins with a broad exploration of the distribution of arrests across various age groups, aiming to discern overarching trends and fluctuations.
+In the data section we clean and process our data.
+Following which in our discussion and results our inquiry begins with a broad exploration of the distribution of arrests across various age groups, aiming to discern overarching trends and fluctuations.
 The age spectrum, a fundamental dimension of this analysis, reveals intriguing patterns that underscore the nuanced nature of arrests in Toronto.
-
-As we embark on this exploration, the data prompts critical questions about the socio-cultural, economic, and systemic factors that contribute to these observed patterns.
-This paper aims to contribute to a nuanced understanding of the dynamics of arrests in Toronto, providing insights that can inform targeted interventions, policy considerations, and community engagement efforts.
+This paper aims to contribute to understanding the dynamics of arrests in Toronto, providing insights that can inform targeted interventions, policy considerations, and community engagement efforts.
 
@@
 However, for the analysis in this paper, we narrowed our focus to specific variables, including Arrest Year, Perceived Race, Age Group, Youth at Arrest, Booked, Search Reason Cause Injury, Search Reason Assist Escape, Search Reason Possess Weapons, and Search Reason Possess Evidence.
-Upon further examination the raw data contained values such as 'NA' under the Booked column and 'None' under the Search column.
-Using the `mutate` function [@rDplyr], we replaced the 'NA' and `None` values with 0 and 1 to summarize the data.
-Additionally, we consolidated similar columns, namely Search Reason Cause Injury, Search Reason Assist Escape, Search Reason Possess Weapons, and Search Reason Possess Evidence, into a unified column labeled "Search Reason." The process involved utilizing the `any()` function [@rDplyr], assigning a value of 1 to indicate the presence of any of these search motives.
+Additionally, we consolidated similar columns, namely Search Reason Cause Injury, Search Reason Assist Escape, Search Reason Possess Weapons, and Search Reason Possess Evidence, into a unified column labeled "Search Reason." utilizing the `any()` function [@rDplyr].
 
@@
 
-Table 1 (@tbl-table1) provides a summary of our data, revealing the number of individuals arrested from each prominent ethnic group in the years 2020 and 2021.
+Since our analysis primarily revolves around the distribution of arrests for each ethnic identity, we extracted all ethnic identities from our data.
+These ethnic identities include Black, White, South Asian, East Asian, Middle Eastern, Latin American, Southeast Asian, and Indigenous.
+The data for the years 2020 and 2021 are summarized in Table 1 (a) (@tbl-table1) and Table 1 (b), respectively.
+
+Table 1 (a) (@tbl-table1) provides a summary of our data, revealing the number of individuals arrested from each prominent ethnic group in the years 2020 and 2021.
+Table 2 (b) shows, out of all the people arrested from each ethnic identity, how many were actually booked under a criminal case.
 
@@
 #| label: tbl-table1
-#| tbl-cap: "Arrests by ethnic background for the year 2020, and 2021"
+#| tbl-cap: "Total number of arrests vs Booked for the year 2020, and 2021"
 ## Table 1: Total Number of Arrests
@@
 table1 <- cleaned_data |> 
+  filter(!perceived_race == "None") |>
   group_by(arrest_year, perceived_race) |> 
@@
 
-kable(table1,
-      align = "c",
-      linesep = "",
-      col.names = c("Ethinic and Identity Based Groups", "2020", "2021"),
-      booktabs = TRUE)
+table2 <- cleaned_data |>
+  filter(booked == 1) |>
+  filter(!perceived_race == "None") |>
+  group_by(arrest_year, perceived_race) |>
+  summarise(n = n(), .groups = "drop") |>
+  arrange(desc(n)) |>
+  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)
+
+knitr::kable(
+  list(table1, table2),
+  col.names = c("Ethinic Identity", "2020", "2021"),
+  booktabs = TRUE,
+  )
+
 ```
 
-### Booked after Arrest (Ethnic and Identity-Based Groups for Year 2020 and 2021)
+### Total number of arrests by age group for the year 2020 and 2021
 
-Table 2 presents a data summary indicating the count of individuals from prominent ethnic groups who were "booked" in criminal cases post-arrest in the years 2020 and 2021.
+It would be interesting to explore the age group distribution of arrests.
+Table 2 (@tbl-table2) exactly presents that.
+A data summary indicating the count of different age groups of individuals from prominent ethnic groups in the years 2020 and 2021.
+All the age groups were the following: Ager 17 years and under, Aged 18 to 24 years, Aged 25 to 34 years, Aged 35 to 44 years, Aged 45 to 54 years, Aged 55 to 64 years, Aged 65 years and over.
+
 Due to the extensive nature of the dataset, encompassing nine different age groups across eight ethnicities, summarizing the information for clarity posed a challenge.
@@
 #| echo: false
-#| tbl-cap: "Arrests Booked by ethnic background for the year 2020, and 2021"
-
-## Table 2: Booked After Arrest
-cleaned_data <- read.csv("../data/cleaned_data.csv")
-
-table2 <- cleaned_data |>
-  filter(booked == 1) |>
-  group_by(arrest_year, perceived_race) |>
-  summarise(n = n(), .groups = "drop") |>
-  arrange(desc(n)) |>
-  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)
-
-kable(table2,
-      align = "c",
-      linesep = "",
-      col.names = c("Ethinic and Identity Based Groups", "2020", "2021"),
-      booktabs = TRUE)
-```
-
-### Total number of arrests by age group for the year 2020 and 2021
-
-Table 3 presents a data summary indicating the count of different age groups of individuals from prominent ethnic groups in the years 2020 and 2021.
-Due to the extensive nature of the dataset, encompassing nine different age groups across eight ethnicities, summarizing the information for clarity posed a challenge.
-Consequently, we have opted to showcase a subset of the data specifically focusing on the age group of 25 to 34 years for improved readability.
-
-```{r}
-#| warning: false
-#| message: false
-#| echo: false
+#| label: tbl-table2
+#| tbl-cap: "Arrests Booked by ethinic background for the year 2020, and 2021"
 
@@
 table3 <- cleaned_data %>%
+  filter(!(age_group == "None") & !(perceived_race == "None")) |>
   group_by(arrest_year, perceived_race, age_group) %>%
@@
 kable(table3_small,
-      caption = "Arrests Booked by ethnic background for the year 2020, and 2021",
       align = "c",
@@
 
+## Results
 We would like to acknowledge that opendatatoronto [@rOpenDataToronto] provides the distribution of the population by ethnicity.
@@
 
-## Results
 Now that we have cleaned the data, we can begin to analyze the data.
 The following section will provide a summary of the data, and the trends that emerge from the data.
-Figure 1, Figure 2, Figure 3, Figure 4 was created using `ggplot` [@rGgplot2], displays the information that can be used as compare the trends as described above.
+Figure 1 (@fig-fig1), Figure 2 (@fig-fig2), Figure 3 (@fig-fig3), and Figure 4 (@fig-fig4) was created using `ggplot` [@rGgplot2], displays the information that can be used as compare the trends as described above.
 
 ### Number of Arrests in year 2020-2021
-With reference to Table 1 and the graph plotted the data shows a non-uniform distribution of arrests across ethnic groups, with the Black ethnic group and White group exhibiting the highest numbers of arrests in both 2020 and 2021.
+With reference to Table 1 (@tbl-table1) and the graph (see @fig-fig1) plotted the data shows a non-uniform distribution of arrests across ethnic groups, with the Black ethnic group and White group exhibiting the highest numbers of arrests in both 2020 and 2021.
 Furthermore, the data indicates a decrease in the number of arrests for the top two ethnic groups from 2020 to 2021.
@@
 #| fig.pos: 'htb'
+#| label: fig-fig1
 
@@
 
-### Number of Arrests Booked in year 2020-2021
-Referring to Table 2 and the plotted graph, the data illustrates a consistent distribution of the number of individuals booked after arrests across various ethnic groups.
+Referring to Table 1 (@tbl-table1) for the proportion of arrests that were booked, we plotted a stacked bar graph (see @fig-fig2), with y-axis representing the total percentage.
+The data illustrates a consistent distribution of the number of individuals booked after arrests across various ethnic groups.
 Notably, there is a higher count of individuals booked after arrest compared to those who were arrested but not booked for a criminal case.
@@
 #| fig.pos: 'htb'
+#| label: fig-fig2
 
@@
 ### Number of Arrests by Age Group for the year 2020-2021
-As depicted in Figure 3, the data reveals a non-uniform distribution of arrests across age groups, with the age groups of 25 to 34 years and 35 to 44 years exhibiting the highest number of arrests in both 2020 and 2021.
+As depicted in Figure 3 (see @fig-fig3), the data reveals a non-uniform distribution of arrests across age groups, with the age groups of 25 to 34 years and 35 to 44 years exhibiting the highest number of arrests in both 2020 and 2021.
 Furthermore, there is a notable decrease in the number of arrests for these top two age groups from 2020 to 2021.
 In contrast, a different trend emerges for other age groups, where the number of arrests has increased over the same time period.
-
-Examining Figure 4, which illustrates the dot plot for the number of arrests by age group in the year 2020 and 2021 for the top two ethnic groups, "Black" and "White," it becomes evident that the age group 25 to 34 years represents the highest number of arrests for individuals of Black ethnicity.
-The trend for Black ethnicity indicates a decreasing curve from this age group onward.
-In contrast, for individuals of White ethnicity, the peak in arrests is observed in the age group 35 to 44 years.
-It is noteworthy that from age 45 to 65, the difference in the number of arrests between both ethnicities starts to converge.
 
@@
 #| fig.pos: 'htb'
+#| label: fig-fig3
 # Bar plot
@@
        x = "Perceived Race",
-       y = "Value",
+       y = "Number of People",
        fill = "Year") +
   theme_minimal() +
-  theme(legend.position = "right") + 
+  theme(legend.position = "bottom") + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+Examining Figure 4 (see @fig-fig4), which illustrates the dot plot for the number of arrests by age group in the year 2020 and 2021 for the top two ethnic groups, "Black" and "White," it becomes evident that the age group 25 to 34 years represents the highest number of arrests for individuals of Black ethnicity.
+The trend for Black ethnicity indicates a decreasing curve from this age group onward.
+In contrast, for individuals of White ethnicity, the peak in arrests is observed in the age group 35 to 44 years.
+It is noteworthy that from age 45 to 65, the difference in the number of arrests between both ethnicities starts to converge.
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| eval: true
+#| fig.cap: "Scatter Plot for the number of arrests by Age-Group for the year 2020, and 2021"
+#| fig.width: 10
+#| fig.height: 6
+#| fig.align: 'center'
+#| fig.pos: 'htb'
+#| label: fig-fig4
 
@@
 ggplot(table3_ii, aes(x = age_group, y = value, color = perceived_race)) +
-  geom_point() + 
-  geom_smooth(method = "loess", se = FALSE) +
+  geom_point(size = 3) +
   facet_wrap(~ year, scales = "free_y") +
@@
        x = "Age Group",
-       y = "Value",
-       color = "Perceived Race") +  # Change fill to color
+       y = "Number of People",
+       color = "Perceived Race") +
   theme_minimal() +
-  theme(legend.position = "right") +
+  theme(legend.position = "bottom") +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
 
-```
 ## Discussion

### 2024-01-25 scripts/00-simulate_data.R

@@ -1,19 +1,64 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from OpenData Toronto for Police Race and Identity Based Data Collection: Arrests & Strip Searches
+# Author: Aviral Bhardwaj
+# Date: 24 January 2024
+# Contact: aviral.bhardwaj@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: none
+# Datasets: https://open.toronto.ca/dataset/police-race-and-identity-based-data-collection-arrests-strip-searches/
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+library(dplyr)
 
 #### Simulate data ####
-# [...ADD CODE HERE...]
+set.seed(1008239407)
+num_observation = 100
+# For booked arrests, which is not always 100
+num_observation_small = sample(50:100, 1, replace=TRUE)
+
+# Fictional Ethnic Identities
+sim_identities <- c("Black", "White", "South Asian", "East Asian", "Middle Eastern", "Latin American", "Southeast Asian", "Indigenous")
+
+# Fictional Age Groups
+sim_age_groups <- c("Aged 17 years and under", "Aged 18 to 24 years", "Aged 25 to 34 years", "Aged 35 to 44 years", "Aged 45 to 54 years", "Aged 55 to 64 years", "Aged 65 years and over")
+
+simulated_arrests_data <-
+  tibble(
+    "arrest_year" = sample(2020:2021, num_observation, replace = TRUE),
+    "percieved_race" = sample(sim_identities, num_observation, replace = TRUE),
+  )
+
+simulated_booked_data <-
+  tibble(
+    "arrest_year" = sample(2020:2021, num_observation_small, replace = TRUE),
+    "perceived_race" = sample(sim_identities, num_observation_small, replace = TRUE),
+  )
+
+
+simulated_age_group_data <-
+  tibble(
+    "arrest_year" = sample(2020:2021, num_observation, replace = TRUE),
+    "perceived_race" = sample(sim_identities, num_observation, replace = TRUE),
[…]

### 2024-01-25 scripts/01-download_data.R

@@ -1,26 +1,33 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from OpenData Toronto for Police Race and Identity Based Data Collection: Arrests & Strip Searches
+# Author: Aviral Bhardwaj
+# Date: 24 January 2024
+# Contact: aviral.bhardwaj@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: none
+# Datasets: https://open.toronto.ca/dataset/police-race-and-identity-based-data-collection-arrests-strip-searches/
+
 
 
 #### Workspace setup ####
 library(opendatatoronto)
 library(tidyverse)
-# [...UPDATE THIS...]
+# Get package data
 
-#### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
+# get package
+package <- show_package("police-race-and-identity-based-data-collection-arrests-strip-searches")
+package
 
+# get all resources for this package
+resources <- list_package_resources("police-race-and-identity-based-data-collection-arrests-strip-searches")
 
+# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
 
-#### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
+# load the first datastore resource as a sample
+data <- filter(datastore_resources, row_number()==1) %>% get_resource()
+
+# Save the data in CSV
+write.csv(data, "../data/raw_data.csv")
 
          

### 2024-01-25 scripts/02-data_cleaning.R

@@ -1,44 +1,63 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from OpenData Toronto for Police Race and Identity Based Data Collection: Arrests & Strip Searches
+# Author: Aviral Bhardwaj
+# Date: 24 January 2024
+# Contact: aviral.bhardwaj@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: none
+# Datasets: https://open.toronto.ca/dataset/police-race-and-identity-based-data-collection-arrests-strip-searches/
 
 #### Workspace setup ####
 library(tidyverse)
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+# Reading the data
+data <- read.csv("../data/raw_data.csv")
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+# Adjusting and cleaning the name of the variable
+cleaned_data <- data %>% 
[…]

### 2024-01-25 scripts/03-test_data.R

@@ -1,15 +1,35 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from OpenData Toronto for Police Race and Identity Based Data Collection: Arrests & Strip Searches
+# Author: Aviral Bhardwaj
+# Date: 24 January 2024
+# Contact: aviral.bhardwaj@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: none
+# Datasets: https://open.toronto.ca/dataset/police-race-and-identity-based-data-collection-arrests-strip-searches/
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+
 
 #### Test data ####
+police_arrest_data <- read.csv("../data/cleaned_data.csv")
+
+# Test Check year is numeric 
+police_arrest_data$arrest_year == as.numeric(police_arrest_data$arrest_year)
+
+# Test Check age group is character
+police_arrest_data$age_group == as.character(police_arrest_data$age_group)
+
+# Test check if booked values is between 0 and 1
+booked_values_check <- all(police_arrest_data$booked >= 0 & police_arrest_data$booked <= 1)
+
+# Test check if search reason is between 0 and 1
+search_reason_check <- all(police_arrest_data$search_reason >= 0 & police_arrest_data$search_reason <= 1)
+
+# Test check if year is between 2020 and 2021
+police_arrest_data$arrest_year <- police_arrest_data %>%
+  filter(arrest_year >= 2020 & arrest_year <= 2021)
+
+# Check if total observations are 32000
+police_arrest_data |> nrow() == 32000

### 2024-01-25 scripts/02-data_cleaning.R
Path Cleaning
- Fixed paths

@@ -12,7 +12,7 @@ library(tidyverse)
 
 #### Clean data ####
 # Reading the data
-data <- read.csv("../data/raw_data.csv")
+data <- read.csv("../inputs/data/raw_data.csv")
 
 # Adjusting and cleaning the name of the variable
 cleaned_data <- data %>% 
@@ -60,4 +60,4 @@ cleaned_data <- cleaned_data |>
   select(
     arrest_year, perceived_race, age_group, youth_at_arrest_under_18_years, booked, search_reason)
 
-write.csv(cleaned_data, "../data/cleaned_data.csv")
+write.csv("../outputs/data/cleaned_data.csv")

### 2024-01-25 scripts/03-test_data.R

@@ -13,7 +13,7 @@ library(tidyverse)
 
 
 #### Test data ####
-police_arrest_data <- read.csv("../data/cleaned_data.csv")
+police_arrest_data <- read.csv("../outputs/data/cleaned_data.csv")
 
 # Test Check year is numeric 
 police_arrest_data$arrest_year == as.numeric(police_arrest_data$arrest_year)

### 2024-01-25 outputs/paper/paper.qmd
Path Cleaning
- Fixed paths
Final Paper Push
Final Paper Push

@@
 author: "Aviral Bhardwaj"
-abstract: The ethnic identity of citizens arrested by the police has consistently been a topic of discussion, often highlighting the pronounced skewness of certain ethnic identities.
-This paper seeks to analyze data from the Toronto Open Data Portal to investigate whether there is any bias in arrests made by the Toronto Police Service between 2020 and 2021, while accounting for factors such as age, booked after arrest, and search on arrest.
-While the population distribution of various ethnic identities varies significantly, it was found that .....
-Further analysis can provide valuable insights for policymakers, law enforcement agencies, and community advocates in fostering a more equitable society while emphasizing the importance of promoting transparency and accountability in policing.
+abstract: The ethnic identity of citizens arrested by police has consistently sparked discussion, highlighting skewness in certain ethnicities.
+This paper analyzes Toronto Police Service arrest data from 2020 to 2021, considering factors such as age and search on arrest.
+While population distribution varies significantly, findings suggest no inherent bias in arrests.
+However, strict enforcement is consistent across all ethnicities and races, even when considering evidence of higher crime rates in specific groups.
+Further analysis can provide insights for policymakers, law enforcement, and community advocates, emphasizing the need for transparency and accountability in policing.
 date: today
@@
 # Save the data in CSV
-write.csv(data, "../data/raw_data.csv")
+write.csv(data, "../../inputs/data/raw_data.csv")
 ```
@@
 # Reading the data
-data <- read.csv("../data/raw_data.csv")
+data <- read.csv("../../inputs/data/raw_data.csv")
 
@@
 
+The analysis presented in this paper aimed to shed light on the potential bias in arrests made by the Toronto Police Service, specifically focusing on ethnic and identity-based groups.
+The findings from the data analysis provide valuable insights into the dynamics of law enforcement activities in the city.
+
+### Lack of Inherent Bias in Arrests
+
+Our analysis suggests no inherent bias in arrests made by the Toronto Police Service.
+The distribution of arrests across various ethnic and identity-based groups does not show a consistent pattern of discrimination.
+This observation is crucial, as it challenges the narrative surrounding racial bias in policing and suggests that strict enforcement is consistent across all ethnicities.
+
+### Age Group Dynamics
+
+The examination of age group dynamics in arrests provides additional nuances to the analysis.
+The age groups of 25 to 34 years and 35 to 44 years emerge as having the highest number of arrests in both 2020 and 2021.
+Interestingly, there is a noticeable decrease in arrests for these age groups from 2020 to 2021, suggesting potential shifts in law enforcement patterns.
+
+The detailed exploration of age groups within specific ethnicities, such as the comparison between "Black" and "White" ethnicities, reveals intriguing patterns.
+The peak in arrests for individuals of Black ethnicity occurs in the age group of 25 to 34 years, while for individuals of White ethnicity, the peak is observed in the age group of 35 to 44 years.
+This divergence in age-specific arrest patterns highlights the importance of considering intersectionality in understanding policing dynamics.
+
+### Implications for Policymaking and Community Engagement
+
+While the analysis in this paper provides valuable insights, it is essential to acknowledge its limitations.
+Further research and a more detailed comparison of specific ethnic groups are necessary for a comprehensive understanding of the data.
+Policymakers, law enforcement agencies, and community advocates can benefit from these insights by considering targeted interventions, policy adjustments, and community engagement efforts.
+The emphasis on transparency and accountability in policing remains crucial, as highlighted in report by the Toronto Police Services Board [@rTPSB2019].
+
+In conclusion, this analysis provides a foundation for informed discussions on policing practices in Toronto.
+It is a call for continued scrutiny, collaboration, and evidence-based policymaking to ensure fair and equitable law enforcement practices in this diverse and dynamic city.
+
 ## References
